---
title: "20180626HomeWork"
output: html_notebook
---

```{r}
p <- 0.95
q <- 0.1
r <- 0.1
M <- matrix(c(p, 1-p, 0, 0, q,1-q,1-r,0,r), 3,3)
init.S <- c(0.9, 0.1,0)
n <- 10
SS <- matrix(0, n, 3)
SS[1,] <- init.S 
for(i in 2:n){
  SS[i, ] <- M %*% SS[i-1,]
}
SS
matplot(SS, type='l')
n <- 10
for(i in 2:n){
  SS[i, ] <- M %*% SS[i-1,]
}
```

```{r}
el <- matrix(c('H','F','F','R','R','H','H','H','F','F','R','R'), 6,2, byrow=TRUE)
el
library(igraph)
g5 <- graph.edgelist(el)
plot(g5)
V(g5)
E(g5)
# t = 5の時のサイズをみる
V(g5)$size <- 100*SS[10,]
p <- 0.95
q <- 0.1
r <- 0.1
E(g5)$width <- 10*c((1-p), (1-q), (1-r), p, q, r)
plot(g5)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

