---
title: "2018_0626_assignment_Takaya"
author: "Hisamitsu Takaya"
date: "2018/6/28"
output: html_document
---

## Assume a population has three statuses, A,B, and C. Set a transition matrix M that moves some fraction of each status to other statuses. Simulate the change of fraction per status.


```{r}
# Set an initial status as A:0.65, B:0.25, C:0.10
A <- 0.65
B <- 0.25
C <- 0.10
S0 <- matrix(c(A,B,C),3,1)

# Set a transition matrix M
M <- matrix(c(0.65,0.3,0.05,0,0.8,0.2,0.6,0.05,0.35),3,3)

# calculate after n-times transition
n <- 30
Sn <- matrix(0,n,3)
Sn[1,] <- S0
for(i in 1:(n-1)){
  Sn[i+1,] <- M %*% Sn[i,]
}
matplot(Sn,type="l")

# difference of fraction before and after transition converge to zero
dif <- Sn[1:(n-1),] - Sn[2:n,]
matplot(dif,type="l")

```

```{r}
# change the initial status at random using rmultinom function
frac <- rmultinom(1,100,abs(rnorm(3)))/100
S0r <- frac
S0r

# calculate after n-times transition
n <- 10
Snr <- matrix(0,n,3)
Snr[1,] <- S0r
for(i in 1:(n-1)){
  Snr[i+1,] <- M %*% Snr[i,]
}
matplot(Snr,type="l")

# any initial status converge to the same fraction

n_iter <- 100
init.stock <- matrix(0,n_iter,3)
after.stock <- matrix(0,n_iter,3)

for(i in 1:n_iter){
  S0r <- rmultinom(1,100,abs(rnorm(3)))/100
  n <- 30
  Snr <- matrix(0,n,3)
  Snr[1,] <- S0r
  for(j in 1:(n-1)){
    Snr[j+1,] <- M %*% Snr[j,]
  }
  init.stock[i,] <- S0r
  after.stock[i,] <- Snr[n,]
}

init <- as.vector(t(init.stock))
after <- as.vector(t(after.stock))
matplot(t(cbind(init,after)),col=1:3,lty=1:3,type="l")

```


## Draw a graph whose nodes are statuses and edges are transition among them. Make the size of vertices proportional to the fraction in the population at a particular point. Make the width of edges proportional to the probability in transition.

```{r}
library(igraph)

# edgelist of initial status
el.i <- matrix(c("A-i","B-i","A-i","C-i","B-i","C-i","C-i","A-i","C-i","B-i","A-i","A-i","B-i","B-i","C-i","C-i"),byrow=T,ncol=2)

# edgelist of after convergency
el.c <- matrix(c("A-c","B-c","A-c","C-c","B-c","C-c","C-c","A-c","C-c","B-c","A-c","A-c","B-c","B-c","C-c","C-c"),byrow=T,ncol=2)

g.i <- graph.edgelist(el.i,directed=T)
g.c <- graph.edgelist(el.c,directed=T)
weight <- c(0.3,0.05,0.8,0.6,0.05,0.65,0.2,0.35)
E(g.i)$weight <- weight

# change the edge color by transition probability
color <- rep("black",8)
color[which(weight>0.5)]<- "red"
color[which(weight<0.1)]<- "blue"

# change the vertex size by status probability
size.i <- S0
size.c <- Snr[n,]
V(g.i)$weight <- size.i
V(g.c)$weight <- size.c

# draw a graph at initial status
plot(g.i,edge.width=E(g.i)$weight*5,edge.curved=0.3,vertex.size=V(g.i)$weight*150,vertex.color=2:4,edge.color=color)

# draw a graph at convergency
plot(g.c,edge.curved=0.3,vertex.size=V(g.c)$weight*150,vertex.color=2:4,edge.color="black")
```


