---
title: "20180703 Classnote"
output: html_notebook
---

```{r}
library(igraph)
# A: adjacency matrix
A <- matrix(c(0,1,1,0,1,0,1,0,1,1,0,1,0,0,1,0), 4,4)
g <- graph.adjacency(A, mode='undirected')
plot(g)
# D: degree matrix
D <- diag(degree(g))
# Laplacian matrix
L <- D - A
range(L - t(L)) # L is symmetric
eigen.out <- eigen(L)
eigen.out[[1]]
round(eigen.out[[1]], 5)
length(which((eigen.out[[1]])^2 < 10^(-16)))
```
wikiより
The number of connected components in the graph is the dimension of the nullspace of the Laplacian and the algebraic multiplicity of the 0 eigenvalue.
つまりconnectされたvertexの数が固有値の中のゼロの数となる
```{r}
M <- matrix(rnorm(100), 10, 10)
M <- M + t(M) 
A <- (M > 1) + 0
diag(A) <- 0
g <- graph.adjacency(A, mode='undirected')
plot(g)
D <- diag(degree(g))
L <- D - A
eigen(L)
help(package=igraph)
clu <- components(g)
groups(clu)
clu
m <- mst(g)
plot(m)
```
Together with Kirchhoff's theorem, it can be used to calculate the number of spanning trees for a given graph. 
spanning trees = 最少のedgeで全てのvertexを結ぶことができるグラフ

応用について
https://web.stanford.edu/class/archive/cs/cs106b/cs106b.1174/lectures/25/Small25.pdf
