---
title: "Mioshibata0425"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R skills
This functions show mean(), var(), skewness(), and kurtosis() of the vector of input.
Then I generated a random values following Normal distribution and applied these functions to the random values.

```{r}
mymean <- function(x,p){
  n<-length(x)
  m<-mean(x)
  ret<-sum((x-m)^p)/n
  return(ret)
}

myvar<-function(x,p){
  n<-length(x)
  m<-mean(x)
  ret<-sum((x-m)^p)/n
  return(ret)
}


myskew<-function(x,p){
  n<-length(x)
  m<-mean(x)
  sd<-sd(x)
  ret<-sum((x-m)^p/sd^p)/n
  return(ret)
}

#Kurtosis of normal ditribution is 3, then we need to substract 3 from the result which is given by the formula to compare the difference from normal distribution.
mykurtosis<-function(x,p){
  n<-length(x)
  m<-mean(x)
  sd<-sd(x)
  ret<-sum((x-m)^p/sd^p)/n -3 
  return(ret)
}

```

Generate a random values in Normal distribution and apply the function above to the value set. 
```{r}
ramdom <-rnorm(100000)

x<-ramdom
```

Each value from functions that I generated.
```{r}
mymean(x,1)
myvar(x,2)
myskew(x,3)
mykurtosis(x,4)
```

Skewness and kurtosis should be essentially zero but the values are not zero because I generated a random value containg 100000 samples. If I generated infinite samples in normal distribution, skewness and kurtosis would be absolutely zero.

##Draw a histogram of the random values.
```{r}
hist(x)
```


###Generate two sets of random values, each of which follow Normal distribution with different mean and sd, respectively.
Combine the two sets and treat them as one set. Draw the histogram of the combined set. Then I applied my functions. 
```{r}
a<-as.vector(rnorm(100000,mean = 0, sd = 1))
b<-as.vector(rnorm(100000,mean = 5, sd = 2))

c<- rbind(a,b)
hist(c)

myskew(c,3)
mykurtosis(c,4)

```

