---
title: "Calculation with matrices 行列で計算"
author: "ryamada"
date: "2016年12月18日"
output: 
  html_document:
    toc: true
    toc_depth: 6
    number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# +, -, x and /
## Calculation of real numbers 実数の計算

```{r}
x <- 3
y <- -5
x+y
x-y
x*y
x/y
```

### Important numbers 大事な値

```{r}
zero <- 0
one <- 1
x
x + zero
x * zero
x * one
x.inverse <- 1/x
x * x.inverse
```
```{r}
x/y
y.inverse <- 1/y
x * y.inverse
```
## Square matrices 正方行列

### $2 \times 2$ matrices
```{r}
X <- matrix(c(1,2,3,4),2,2)
X
```
```{r}
Y <- matrix(c(4,5,6,7),2,2)
Y
X+Y
X-Y
X %*% Y
Y %*% X
```

Calculate yourself without computers. 計算機を使わずに計算すること。

#### Important matrices

```{r}
ONE <- matrix(c(1,0,0,1),2,2)
ONE
ZERO <- matrix(0,2,2)
ZERO
X.inverse <- solve(X)
X
X.inverse

X %*% X.inverse
X.inverse %*% X
```

```{r}
# x/y
Y.inverse <- solve(Y)
X %*% Y.inverse
Y.inverse %*% X
```

### $n \times n$ matrices $n \times n$ 行列

```{r}
n <- sample(3:5,1)
X <- matrix(sample((-20):20,n^2,replace=TRUE),n,n)
Y <- matrix(sample((-20):20,n^2,replace=TRUE),n,n)
X
X + Y
X - Y
X %*% Y
X %*% solve(X)
solve(X) %*% X
X %*% solve(Y)
solve(Y) %*% X
```

#### Important matrices
```{r}
ONE <- diag(rep(1,n))
ONE
ZERO <- matrix(0,n,n)
ZERO
X.inverse <- solve(X)
X
X.inverse
```

# Power and exponential

## Power to integer
```{r}
x <- 3
p <- 2
x^p
x * x
p <- 3
x^3
x * x * x
```

```{r}
X <- matrix(c(1,2,3,4),2,2)
p <- 2
X %*% X
p <- 3
X %*% X %*% X
```

## Power to non-integer
```{r}
x <- 2
p <- 2
x. <- x^p
x.
p.inverse <- 1/p
p.inverse
x.^p.inverse
```

## Power of diagonal matrix
```{r}
X <- matrix(c(2,0,0,3),2,2) # Diagonal matrix
X
diag(c(2,3))
X %*% X
p <- 2
p.inverse <- 1/p
diag.X <- diag(X)
diag.X
diag(diag.X^p)
diag(diag.X^p.inverse)
```

## Decomposition of matrix 行列の分解

Eigen value decomposition　固有値分解

$$
X = V S V^{-1}
$$
$$
X^2 = (V S V^{-1}) (V S V^{-1}) = (VS)(V^{-1}V)(SV^{-1})=V(S^2)V^{-1}
$$
$$
X^p = V(S^p)V^{-1}
$$
```{r}
X <- matrix(c(2,3,4,5),2,2)
eigen.out <- eigen(X)
eigen.out
V <- eigen.out[[2]]
S <- diag(eigen.out[[1]])
V %*% S %*% solve(V)
X %*% X
V %*% diag(eigen.out[[1]]^2) %*% solve(V)
X. <- V %*% diag(eigen.out[[1]]^(1/2)) %*% solve(V)
X.
s <- complex(real=eigen.out[[1]],imaginary=0)
s^(1/2)
X.. <- V %*% diag(s^(1/2)) %*% solve(V)
X..
X.. %*% X..
```

## Matrix exponential

指数関数の性質 Exponential function behaves:

$$
\frac{d}{dx} e^x= e^x
$$
$$e^0=1$$

指数関数の定義式

Definition of exponential function.

$$
e^x = \sum_{k=0}^{\infty} \frac{1}{k!}x^k
$$


$$
e^X = \sum_{k=0}^{\infty} \frac{1}{k!}X^k
$$

$$
e^{VSV^{-1}} = \sum_{k=0}^{\infty} \frac{1}{k!}VS^kV^{-1} = V(\sum_{k=0}^{\infty} \frac{1}{k!} S^k) V^{-1}
$$

対角成分ごとに$e^x$を計算できる

You only have to calculate $e^x$ of each diagonal element.

```{r}
X <- matrix(c(2,3,4,5),2,2)
eigen.out <- eigen(X)
s <- eigen.out[[1]]
V <- eigen.out[[2]]
s
V

V %*% diag(exp(s))%*% solve(V)
```

# 非正方行列の積 Multiplication of non-square matrices 

$M_{n,m}$, $n \times m$ matrix can be multiplied by $M_{k,n}$ from its left and by $M_{m,k}$ from its right. The dimension of the products are $k\times n$ and $m\times k$, respectively.

$n \times m$ 行列 $M_{n,m}$は、左から$M_{l,n}$行列を掛けることができ、右から$M_{m,k}$行列を掛けることができる。生じる行列はそれぞれ、$k \times n$、$m \times k$ である。

```{r}
n <- 4
m <- 3
k <- 2

M1 <- matrix(1:(n*m),n,m)
M1
dim(M1)

M2 <- matrix(1:(k*n),k,n)
M2
dim(M2)
M21 <- M2 %*% M1
M21
dim(M21)

M3 <- matrix(1:(m*k),m,k)
M13 <- M1 %*% M3
M13
dim(M13)
```

## ベクトルの内積 Inner product of vectors

A vector with n elements can be considered $n\times 1$ matrix or $1 \times n$ matrix.

要素数 n のベクトルは$n\times 1$行列、$1 \times n$行列とみなせる。

```{r}
n <- 3
v1 <- c(3,5,6)
v2 <- c(1,2,4)
V1 <- matrix(v1,nrow=n)
V1
V2 <- matrix(v2,nrow=1)
V2
```

$n \times 1$ matrix can be multiplied by $1 \times n$ matrix from its left. The product is $1 \times 1$ matrix, or a scaler value. 

$n \times 1$ 行列には、その左側から$1 \times n$行列を掛けることができる。その積は$1 \times 1 $ 行列、もしくは、スカラー値である。

```{r}
V2 %*% V1
```

The following returns the same value. 次のような計算と同じである。

```{r}
sum(v1*v2)
```

# Differential equation 微分方程式

## Single differential equation １つの微分方程式
$$
\frac{d x(t)}{dt} = a x
$$

解
$$
x(t) = b \times e^at
$$

## System of differential equation 連立微分方程式

$$
\begin{pmatrix} \frac{dx_1}{dt} \\ \frac{dx_2}{dt} \end{pmatrix} = \begin{pmatrix} a & b \\ c & d \end{pmatrix} \begin{pmatrix} x_1\\ x_2 \end{pmatrix}
$$

The solution of this system is known as,
この連立微分方程式の解は以下であると知られている。


$$
\begin{pmatrix} x_1(t)\\ x_2(t) \end{pmatrix} = e^{t\times A} \begin{pmatrix} x_1(0)\\ x_2(0) \end{pmatrix}\\
A = V S V^{-1}\\
S = diag(e^{\lambda_A})\\
\Sigma(t) = diag(e^{t \times \lambda_A})\\
e^{t \times A} = V \Sigma(t) V^{-1}
$$
```{r}
A <- matrix(c(0.5,1,-1,-0.3),2,2)
A
eigen.out <- eigen(A)
S <- diag(eigen.out[[1]])
V <- eigen.out[[2]]

t <- seq(from=0,to=30,length=100)
x0 <- c(5,-1)
xs <- matrix(0,length(t),2)
for(i in 1:length(t)){
  etA <- V %*% diag(exp(t[i]*diag(S))) %*% solve(V)
  xs[i,] <- etA %*% x0
}
matplot(xs,type="l")
plot(Re(xs),type="l") # state space plot
```

# Exercises

## Exercise 1-1
正方行列の和を計算する関数は次のように作れる。
それにならって、行列の積(非正方行列を含む)を計算する関数を作成、それが正しいことを確かめよ。

The following function calculate sum of matrices. Make functions of production of (non-square) matrices in the similar way.
```{r}
my.matrix.sum <- function(x,y){
  dm <- dim(x)
  ret <- matrix(0,dm[1],dm[2])
  for(i in 1:dm[1]){
    for(j in 1:dm[2]){
      ret[i,j] <- x[i,j]+y[i,j]
    }
  }
  return(ret)
}

x <- matrix(c(2,3,4,5),2,2)
y <- matrix(c(4,5,6,7),2,2)
x + y
my.matrix.sum(x,y)
```

## Exercise 1-2

行列の冪$x^p$を計算する関数を作成せよ

Make a function of power of matrix.

## Exercise 1-3

行列の指数関数$e^{t \times X}$を計算する関数を作成せよ

Make a function of exponential of matix.

## Exercise 1-4

色々な$2 \times 2$ 行列を作り、それに基づく２変量連立微分方程式に関する解をプロットせよ。matplotと状態空間co-plotをせよ。カーブが多様であるように行列を選べ。

Make various $2 \times 2 $ matrices and plot their answers of the system of differential equations of 2 variables. Plot them in two ways (matplot-way and co-plot-way in state space).
Select matrices so that the curves of the variables are heterogeneous.

## Exercise 1-5

$3 \times 3$ 行列による3変量の場合を同様に行え。matplot-wayと3次元状態空間プロットを行え。
さまざまなカーブを描くように、行列を選べ。

Make various $3 \times 3$ matrices for three variables' system of differential equations.

plot them in two ways (matplot-way and 3D-plot-way in state space).

Select matrices so that the curves or the variables are heterogeneous.
