---
title: "2018_0508_Assignment_Takaya"
author: "Hisamitsu Takaya"
date: "2018/5/10"
output: html_document
---

## Assignment

### Report R codes to generate positive-definite matrix, that is the matrix satisfying the features of covariance matrix, i.e., symmetric and all eigenvalues being positive with two ways; (i) by generating data matrix and making its covariance matirix, (ii) by generating a diagonal matrix whose diagonal elements >0 and sandwich it with rotation matrix and its inverse

---

#### (i) Generate positive-definite matrix by generating data matrix and making its covariance matrix

```{r}
# generate data from normal distribution which has some relation

d <- 5
n_sample <- 100
X <- matrix(rep(n_sample*d,0),n_sample,d)
x <- rnorm(n_sample,3,3)
for(i in 1:d){
  if(i == 1){
    X[,i] <- x
  }else{
    X[,i] <- x + rnorm(n_sample,0,3)
  }
}

# make covariance matrix
M <- cov(X)
M
```

```{r}
# check covariance matrix has the features of positive-definite matrix

M - t(M) # should be zero matrix

eigen.M <- eigen(M)
eigen.val <- eigen.M[[1]]
eigen.vec <- eigen.M[[2]]

eigen.val > 0 # all values should be TRUE
round(t(eigen.vec) %*% eigen.vec,8) # should be identity matrix
```

positive-definite matrix meet the following condition 

$$
v^TMv>0(v \neq \bf{0})
$$

```{r}

n_iter <- 1000
v <- matrix(rep(d*n_iter,0),n_iter,d)
s <- rep(n_iter,0)
for(i in 1:1000){
  v[i,] <- rnorm(d)
  s[i] <- t(v[i,]) %*% M %*% v[i,]
}
plot(seq(n_iter),s,ylim=c(0,max(s)))
abline(h=0,col=2)
range(s)

```

#### (ii) Generate positive-definite matrix by generating a diagonal matrix whose diagonal elements >0 and sandwich it with rotation matrix and its inverse

```{r}
# generate a diagonal matrix
D <- diag(runif(d,1,5))
D
```

To generate symmetric matrix, rotation matrix must be orthogonal matrix. A orthogonal matrix can be made eigen vectors of a symmetric matrix.

```{r}
# make a rotation matrix
R <- matrix(rnorm(d^2),d,d)
R2 <- eigen(R+t(R))[[2]]
R2
Y <- R2 %*% D %*% solve(R2)
Y

round(Y - t(Y),8)
```

```{r}
# matrix Y has the same eigen values with matrix D

eigen(D)[[1]]
eigen(Y)[[1]]

# check matrix Y2 has the features ofpositive-definite matrix

n_iter <- 1000
v <- matrix(rep(d*n_iter,0),n_iter,d)
s <- rep(n_iter,0)
for(i in 1:1000){
  v[i,] <- rnorm(d)
  s[i] <- t(v[i,]) %*% Y %*% v[i,]
}
plot(seq(n_iter),s,ylim=c(0,max(s)))
abline(h=0,col=2)
range(s)

```

