---
title: "Assignments 0508"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Positive definite matrix is defined that matrix is symmetric and all its eigenvalules are positive.
This is the one of the example.

$$
A=
\begin{pmatrix}
 2 & 1\\
 1 & 2
\end{pmatrix}
$$
$$
det(A)= 4-1=3>0\\
$$

Assume $\lambda$ as the eigenvalues of matrix of A,

$$
(2-\lambda)^2 -1 = 0\\
(\lambda-3)(\lambda-1)=0\\
\lambda=3,1 >0
$$
Both eigne values are positive.
Then the matrix of A can be said as positive definite matrix.

###Generating the positive definite 2 $\times$ 2matrix.
Assume the positive definite matrix as M,

$$
M=
\begin{pmatrix}
 a & b\\
 b & a
\end{pmatrix}
$$
The eigen value is $\lambda$,
$$
(a-\lambda)^2 - b^2=0\\
(\lambda-a+b)(\lambda-a-b)=0\\
\lambda= a-b, a+b
$$
Then if $a-b>0 and a+b>0$, the matrix of M is positive definite matrix.

###R code,
 (i) by generating data matrix and making its covariance matirix,

```{r}
k <- 3
n <- 100
X <-matrix(rnorm(k^2),k,k)
M <- cov(X)
M
```
Showing that eigen values are positive
```{r}
eigen(M)
```

(ii) by generating a diagonal matrix(D) whose diagonal elements >0 and sandwich it with rotation matrix(R) and its inverse wiht this formula,
$R^{-1}MR=D$

Generating the positive definiti matrix of M
```{r}
k <- 3
e <- runif(k,0)
D<-diag(e)

##Generating rotation matrix and its inverse 
library(GPArotation)
R <- Random.Start(k)
M <- R %*% D %*% t(R)
M
```

Showing the eigenvalue
```{r}
eigen(M)
```


