---
title: "R Notebook"
output: html_notebook
---

```{r}
N <- 100
p <- 0.7

AA <- N * p^2
Aa <- N * 2 * p * (1-p)
aa <- N * (1-p)^2

Genotype <- matrix(rep(0,3*31), 31, 3)

for (k in 1:31){
  Genotype[k,1] <- k + 39
  Genotype[k,2] <- 62 - 2 * k
  Genotype[k,3] <- k - 1
}
Genotype

install.packages('HardyWeinberg')
library('HardyWeinberg')

HWE.results <- rep(0, 31)
HWExact(Genotype[1,])$pval

for (k in 1:31){
  HWE.results[k] <- HWExact(Genotype[k,])$pval
}
HWE.results
```

###packageを用いず、山田先生のtextより
[こちら](http://www.statgenet.med.kyoto-u.ac.jp/StatGenet/ryamada_bon/Whole/Whole20100523/ryamadaBonEn_Whole.html)y取り
```{r}
hweExact<-function(g=c(813,182,5)){ # number of people with three genotypes
 n<-sum(g) # total 
 nA<-2*g[1]+g[2] # No. allele A
 na<-2*g[3]+g[2] # No. allele a
 evod<-g[2]%%2 # Heterozygoutes are even or odd
 maxAa<-min(nA,na)-evod
 Aa<-seq(from=evod,to=maxAa,by=2) # list of possible heterozygous people number
 AA<-(nA-Aa)/2 # corresponding No. AA
 aa<-(na-Aa)/2 # corresponding No. aa
 obs<-(g[2]-evod)/2+1 # observed heterozygoutes
 prob<-rep(0,length(Aa)) # exact probs of all cases
 prob<-exp(n*lgamma(2+1)+lgamma(nA+1)+lgamma(na+1)-lgamma(2*n+1)-(AA*lgamma(2+1)+Aa*lgamma(1+1)+aa*lgamma(2+1))+lgamma(n+1)-(lgamma(AA+1)+lgamma(Aa+1)+lgamma(aa+1)))
 p.value<-sum(prob[prob<=prob[obs]]) # p-value
 # Aa No. heteros
 # prob exact probs
 # obsprob exact prob of observed table
 # p.value 
 list(Aa=Aa,prob=prob,obsprob=prob[obs],p.value=p.value)
}
xx<-hweExact(c(813,182,5))
xx$p.value # test p-value
sum(xx$prob) # sum of all exact probs is 1
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

