---
title: "Assignment0528"
output: html_document
---


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
library(rgl)
knit_hooks$set(rgl = hook_webgl)
library(matlib)
knit_hooks$set(matlib = hook_webgl)
library(car)
library(matlib)
```

$y=\beta_1x_1 + \beta_2x_2$ is a plane in 3d space where $x_1,x_2,y$ are three axes.
Given $(\beta_1, \beta_2) = (2,-1)$ in 3d space.
```{r}
K <- 200
x1 <- matrix(rnorm(K),ncol=1)
x2 <- matrix(rnorm(K),ncol=1)
y <- 2*x1-x2
p2<-cbind.data.frame(x1,x2,y)
fit <- lm(y ~ x1 + x2, p2)
coefs <- coef(fit)
```
I coudn't draw a plane directrly from the vector..
Instead I draw scatter plot and draw a surface which is fitted to the scatter plot.
That surface is the plane that I want to draw.
```{r rgl=TRUE}
plot3d(p2)
planes3d(coefs[2], coefs[3], -1, coefs[1], col="blue", alpha=0.5)
```


When two matrixes are given, and we use at least squares to calculate $\hat\beta$.
$\hat\beta=(X^TX)^{-1}X^Ty $
Calucalating the $\beta$ from that formula,
```{r}
y<-matrix(c(1,2,3,4),ncol=1)
X<-matrix(c(0,1,0,1,0,0,1,1),ncol=2)
y
X
beta <- solve(t(X)%*%X)%*%t(X)%*%y
beta
```
But the result is a coordinate so I didn't get the curved surface over grid points of $(\beta_1, \beta_2)$.

```{r}
plot(beta)
```

Given,$y=\beta_1h_1(x) + \beta_2h_2(x)$ and $(h_1(x), h_2(x))=(x_1+x_2, x_1^2)$,I drew the curved surface in 3d space.
I couldn't draw the curved surface directly then I used function scatter3d with fit =  "smooth" option that can drew the curved surface.  
```{r}
K <- 200
x1 <- matrix(rnorm(K),ncol=1)
x2 <- matrix(rnorm(K),ncol=1)
h1<-x1+x2
h2<-x1^2
y <- 2*h1-h2


```

```{r rgl=TRUE}
scatter3d(x1,x2,y,fit = "smooth",grid = FALSE)
